\Title

### 2021-02-04 manuscript/content.tex
Edits to introduction

Several methods exist in the literature for interpolation in two dimensions,
for example continuous curvature splines in tension \citep{smith1990},
bi-harmonic (thin-plate) splines \citep{sandwell1987}, and kriging \citep{hansen1993}.
These general-purpose methods have limitations when it comes to interpolating
potential field data:
(i) they are not able to take into account the variable height of the
observation points and
(ii) the interpolating functions are not necessarily harmonic functions, which
is the underlying assumption behind many processing techniques
(e.g., upward continuation and vertical derivatives).

A widely used method for interpolating gravity and magnetic data
is the equivalent sources technique (also known as equivalent layer, radial
basis functions, or Green's functions interpolation), first introduced by
\citet{dampney1969}.
It consists in fitting a model of elementary sources to the data and using this
model to predict new data values.
Besides interpolation, equivalent sources have been used for
reduction-to-the-pole of magnetic data
\citep{silva1986, nakatsuka2006, guspi2009}, upward
continuation \citep{emilia1973, li2010}, joint processing of gravity gradient
data \citep{barnes2011}, modelling the lithospheric magnetic field
\citep{kother2015}, recovering the magnetic induction vector from
total-field magnetic anomalies \citep{li2020}, and more.

Many variants of the equivalent sources technique have been proposed, often
attempting to obtain faster or more accurate solutions.
The key factors that vary between them are: (i) the type of source, (ii)
the location of the sources, and (iii) the solution strategy.
The type of source is most commonly a point mass for gravity or dipole for
magnetics \citep[e.g.,~][]{vonfrese1981, silva1986, mendonca1994, siqueira2017}.
However, right-rectangular prisms \citep[e.g.,][]{barnes2011, jirigalatu2019,
li2020} and tesseroids \citep{bouman2016} have also been used successfully.
In fact, even point sources with a simple inverse distance function, instead of
actual gravity or magnetic fields, can be used as
equivalent sources \citep{cordell1992}.
The sources are often distributed on a regular grid at a constant depth
\citep[e.g.,~][]{leao1989, barnes2011, oliveira2013}
or placed beneath each data point \citep[e.g.,~][]{cordell1992, siqueira2017}.
The model is usually estimated through damped least-squares, which imposes a
heavy computational load when the number of data points is large (e.g.,
airborne and satellite surveys).
To reduce the computational load, \citet{mendonca1994} built the solution
iteratively by incorporating one data point at a time using the ``equivalent
data concept''.
\citet{leao1989} processed the input data using a moving window, only fitting the
data inside the window and predicting observations at the center of the window.
\citet{li2010} and \citet{barnes2011} apply different operations to generate a
sparse representation of the sensitive matrix (respectively, wavelet
compression and quadtree discretization), which significantly improves the
speed of the least-squares solution.
\citet{oliveira2013} parametrized the equivalent layer as a piecewise bivariate
polynomial function, reducing the number of parameters in the solution.
\citet{siqueira2017} developed an iterative solution in which the sensitivity
matrix is transformed into a diagonal matrix with constant terms through the
``excess mass criterion''.
\citet{jirigalatu2019} applied the Gauss-FFT method to speed up the forward
modelling operations and solved the least-squares problem using steepest
descent to avoid calculating the Hessian matrix and solving linear systems.

Many of the existing methods solve under-determined problems, requiring a much
larger number of equivalent sources than the number of data points.
Some achieve greater efficiency by restricting their applications
specific data types \citep{siqueira2017},
interpolating only on regular grids \citep{leao1989},
or requiring already gridded data \citep{takahashi2020},
to name a few.
Furthermore, many of the optimizations proposed are also complex to implement
in a computer program, limiting their wider adoption.

@@

Through tests on synthetic data, we show that: (i) the \emph{block-averaged}
sources are able to achieve the same accuracy as other traditional equivalent
source layouts while using a fraction of the number of sources, and (ii) the
\emph{gradient boosting} algorithm greatly reduces the computational memory
required to fit very large datasets without sacrificing prediction accuracy.
Finally, a combination of both strategies are used to process a collection of
around 1.7 million ground gravity data measurements from Australia.

>>>>

Several methods exist in the literature for interpolation in two dimensions,
for example continuous curvature splines in tension \citep{smith1990},
bi-harmonic (thin-plate) splines \citep{sandwell1987}, and kriging \citep{hansen1993}.
These general-purpose methods have limitations when it comes to interpolating
potential field data:
(i)~they are not able to take into account the variable height of the
observation points and
(ii)~the interpolating functions are not necessarily harmonic, which
is the underlying assumption behind many processing techniques
(e.g., upward continuation and vertical derivatives).

A widely used method for interpolating gravity and magnetic data
is the equivalent sources technique (also known as equivalent layer, radial
basis functions, or Green's functions interpolation), first introduced by
\citet{dampney1969}.
It consists in fitting a model of finite elementary sources to the data and
using this model to predict new data values.
Besides interpolation, equivalent sources have been used for
reduction-to-the-pole of magnetic data
\citep{silva1986, nakatsuka2006, guspi2009}, upward
continuation \citep{emilia1973, li2010}, joint processing of gravity gradient
data \citep{barnes2011}, modelling the lithospheric magnetic field
\citep{kother2015}, recovering the magnetic induction vector from
total-field magnetic anomalies \citep{li2020}, and more.

Many variants of the equivalent sources technique have been proposed, often
attempting to obtain faster or more accurate solutions.
The key factors that vary between them are: (i) the type of source, (ii)
the location of the sources, and (iii) the solution strategy.
The type of source is most commonly a point mass for gravity or dipole for
magnetics \citep[e.g.,~][]{vonfrese1981, silva1986, mendonca1994, siqueira2017}.
However, right-rectangular prisms \citep[e.g.,][]{barnes2011, jirigalatu2019,
li2020} and tesseroids \citep{bouman2016} have also been used successfully.
In fact, even point sources with a simple inverse distance function, instead of
actual gravity or magnetic fields, can be used as
equivalent sources \citep{cordell1992}.
The sources are often distributed on a regular grid at a constant depth
\citep[e.g.,~][]{leao1989, barnes2011, oliveira2013}
or placed beneath each data point \citep[e.g.,~][]{cordell1992, siqueira2017}.
The model is usually estimated through damped least-squares, which imposes a
heavy computational load when the number of data points is large (e.g.,
airborne and satellite surveys).
To reduce the computational load, \citet{mendonca1994} built the solution
iteratively by incorporating one data point at a time using the ``equivalent
data concept''.
\citet{leao1989} processed the input data using a moving window, only fitting the
data inside the window and predicting observations at its center.
\citet{li2010} and \citet{barnes2011} apply different operations to generate a
sparse representation of the sensitive matrix (respectively, wavelet
compression and quadtree discretization), which significantly improves the
speed of the least-squares solution.
\citet{oliveira2013} parametrized the equivalent layer as a piecewise bivariate
polynomial function, reducing the number of parameters in the solution.
\citet{siqueira2017} developed an iterative solution in which the sensitivity
matrix is transformed into a diagonal matrix with constant terms through the
``excess mass criterion''.
\citet{jirigalatu2019} applied the Gauss-FFT method to speed up the forward
modelling operations and solved the least-squares problem using steepest
descent to avoid calculating the Hessian matrix and solving linear systems.

Many of the existing methods solve under-determined problems, requiring a much
larger number of equivalent sources than the number of data points.
Some achieve greater efficiency by restricting their applications
to specific data types \citep{siqueira2017},
interpolating only on regular grids \citep{leao1989},
or requiring already gridded data \citep{takahashi2020},
to name a few.
Furthermore, many of the optimizations proposed are also complex to implement
in a computer program, limiting their wider adoption.

@@

Through tests on synthetic data, we show that:
(i)~the \emph{block-averaged} sources are able to achieve the same accuracy as
other traditional equivalent source layouts while using a fraction of the
number of sources, and
(ii)~the \emph{gradient boosting} algorithm greatly reduces the computational
memory required to fit very large datasets without sacrificing prediction
accuracy.
Finally, a combination of both strategies is used to process a collection of
around 1.7 million ground gravity data measurements from Australia.
